The expected shape of the Milky Way's Dark Matter halo

### 2018-10-12 Paper/paper.tex
edit

In this work we use the results of the state-of-the art Auriga
simulations \cite{auriga}. 
The sample starts with a set of 30 isolated halos from the Evolution
and Assembly of GaLaxies and their Environments (EAGLE) project \cite{Eagle}. 
These halos were randomly selected from a sample of the most isolated
quartile of halos whose virial mass $M_{200}$ (200 times the critical
density of the Universe) varied between $10^{12}M_\odot$ and $2\cdot
10^{12}M_\odot$.
These halos were re-simulated with the moving mesh code AREPO
\citep{arepo} using different degrees of resolution and physics.
All 30 halos were simulated within a level-4-degree of
resolution defined for Aquarius simulations corresponding to $\sim
3\times 10^6$ high resolution particles of mass $\sim 2.5 \times 10^5
M_\odot$. 
%The principal details of each of these halos are consigned
%on the table \ref{tab:level4}. 
From these halos, 6 were re-simulated at a higher resolution (Level-3)
by increasing the number of particles by a factor of $8$.
For each ach halo in Level-3 and Level-4 there are two version: one
only includes DM particles, while the other also includes baryons with
magneto-hydrodynamicl (MHD) physics.
%re-simulated at level 3 (higher) resolution taking into account a
%spatial factor of 2 in each dimension. For more information about
%level 3 halos, their details are printed on table
%\ref{tab:level3}.

@@

There is no trivial way to calculate the DM halo shape at a determined
radius due to the discretization of particles. For practical effects,
observational models focus on the shape of the DM isopotential
\cite{isopotential} or isodensity \cite{isodensity} contours. In
cosmological simulations, it is preferred to work with the isodensity
contours which is directly calculated from particle
positions. However, these density contours are not smooth and are very
sensitive to the presence of small satelites \cite{Springel
  Isodensity}. For this reason we determine the shape taking into
account volume-enclosed particles, rather than shell-enclosed. While
this approach allows the cumulative influence of outter shells on the
outter shapes, it has been demonstrated for DM-only simulations that
this effect becomes important only for big radii \cite{vera-Ciro
  aquarius}.  We therefore follow the method of \cite[Allgood et
  al. 2006]{Allgood_et_al._2006} that consits in the use of the
reduced inertia tensor,

@@

with components weighted by the k-th particle distance $d^2=x^2+y^2+z^2$.\\

The diagonalization of this tensor yields the principal axes of the
structure as well as the eigen-quantities which are proportional to
the squared principal axes $a>b>c$. We start our calculations taking
into account particles within a sphere of radius $R$ and then
recharacterize the triaxial parameters by taking into account
particles within an ellipsoid of semi-axes $r,r/q,r/s$ and weighted
distance $d^2=x^2+(y/q)^2+(z/s)^2$, where ere $q = b/a$ and $s=c/a$
are the previously calculated axial ratios. We repeat this process
until the average deviation of semi-axes is less than $10^{-6}$.

>>>>

@@

There are two main ways to estimate the DM halo shape at a fixed
radius: by computing isopotential or isodensity surfaces.
Observational inference models estimate the shape from the
isopotential contours, while simulations work with the isodensity
contours which can be  directly calculated from particle positions.  
However, the density contours are not smooth and are very 
sensitive to the presence of small satelites.
For this reason we choose to measure the shape by taking
volume-enclosed particles, rather than shell-enclosed.

We follow the shape measurement method presented by
\cite{Allgood_et_al._2006} that consits in the using the reduced
inertia tensor,

@@

with the positions components weighted by the k-th particle distance
$d_k^2=x_k^2+y_k^2+z_k^2$, the particle positions are measured from
the minimum of the gravitational potential in each halo.

The diagonalization of this tensor yields the principal axes of the
structure as well as the eigen-quantities which are proportional to
the squared principal axes $a>b>c$. 
We start our calculations taking
into account particles within a sphere of radius $R$ and then
recharacterize the triaxial parameters by taking into account
particles within an ellipsoid of semi-axes $r,r/q,r/s$ and weighted
distance $d^2=x^2+(y/q)^2+(z/s)^2$, where $q = b/a$ and $s=c/a$
are the previously calculated axial ratios. 
We repeat this process until the average deviation of semi-axes is
less than $10^{-6}$.

This is the same method used to estimate the halo shape in the DM-only
Aquarius simulations \citep{Vera-Ciro_et_al._2011}.
